---
title: "Thesis Data Viz"
author: "Ryan Calhoun & Nicholas Paterna"
date: "6/3/2021"
output: 
  html_document:
    theme: paper
    highlight: zenburn
    center: TRUE
editor_options: 
  
  chunk_output_type: console
---

```{r setup, include=FALSE, fig.align = "center"}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(RColorBrewer)
library(lattice)
library(patchwork)
library(corrplot)
library(MQMF)

if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("Glimma")

library(limma)
library(Glimma)

#setwd("/Users/Caronelab/Desktop/Spicy_Pepper_Project/R_Workflow") #Lab Computer
#setwd("/Users/nickpaterna/Repositories/Spicy_Pepper_Project/R_Workflow") #Nick's MacBook
setwd("/Users/calhoun/Repositories/Spicy_Pepper_Project/R_Workflow") #Ryan's MacBook

```

## Loading in Raw Data
  Here we imported **all** of the raw data before _tidying_ it. All functions were performed utilizing the default **Tidyverse** package!
  
```{r Loading in Data}

genes_fpkm <- read_csv("genes_fpkm.csv") #Reading in FPKM Values
protein_data <- read_csv("Spicy_Pepper_cufflinks_information.csv") #Pepper info

```

## Tidying the Raw Data

```{r Tidying the Data}

tidy_fpkm <- genes_fpkm %>%
  pivot_longer(cols = 4:51,
               names_to = "Sample_ID",
               values_to = "FPKM") #Shrink dataframe width by creating 'Sample ID' col

final_tidy <- tidy_fpkm %>%
  full_join(protein_data,
             by = "Sample_ID") %>%
  rename(c(Locus = locus,
           Gene_ID = genes_fpkm_with_protein_charact)) 
#Add pepper information to tidied dataset, rename cols to make working w data easier

trimmed_tidy <- final_tidy %>%
  mutate(Sample_ID = str_sub(Sample_ID, 1, -5)) %>% #Shrinks sample name
  mutate(Forward = str_detect(Sample_ID, "R1")) %>% #Logic output for direction
  mutate(Direction = case_when(Forward == TRUE ~ "Forward",
                               Forward == FALSE ~ "Reverse")) %>%
  select(-Forward) %>% #Removes logic col now that we have actual direction
  mutate(Sample_ID = str_sub(Sample_ID,1, 5))


splom_object <- genes_fpkm %>%
  pivot_longer(cols = 4:51,
               names_to = "Sample_ID",
               values_to = "FPKM") %>%
  mutate(FPKM = ifelse(FPKM == 0, 1, FPKM), 
         FPKM_log2 = log2(FPKM)) %>%
  mutate(Sample_ID = str_sub(Sample_ID, 1, -5)) %>%
  rename(Gene_ID = genes_fpkm_with_protein_charact) %>%
  select(c(-Protein_Function, -locus, -FPKM)) %>%
  pivot_wider(names_from = Sample_ID,
              values_from = FPKM_log2) %>%
  rename(HabP1_R1 = HabP1_R1,HabP1_R2 = HabP1_R2,JalP1_R1 = JalP1_R1, JalP1_R2 = JalP1_R2,SerP1_R1 = SerP1_R1,SerP1_R2 = SerP1_R2,WaxP1_R1 = WaxP1_R1,WaxP1_R2 = WaxP1_R2,CheP1_R1 = CheP1.LCL3435_L3_R1,CheP1_R2 = CheP1.LCL3435_L3_R2,CheP2_R1 = CheP2.LCL3437_L3_R1,CheP2_R2 = CheP2.LCL3437_L3_R2,CheS1_R1 = CheS1.LCL3434_L3_R1,CheS1_R2 = CheS1.LCL3434_L3_R2,CheS2_R1 = CheS2.LCL3436_L3_R1,CheS2_R2 = CheS2.LCL3436_L3_R2,GhoP1_R1 = GhoP1.LCL3453_L3_R1,GhoP1_R2 = GhoP1.LCL3453_L3_R2,GhoS1_R1 = GhoS1.LCL3452_L3_R1,GhoS2_R2 = GhoS1.LCL3452_L3_R2,HabP2_R1 = HabP2.LCL3451_L3_R1,HabP2_R2 = HabP2.LCL3451_L3_R2,HabS1_R1 = HabS1.LCL3448_L3_R1,HabS1_R2 = HabS1.LCL3448_L3_R2,HabS2_R1 = HabS2.LCL3450_L3_R1,HabS2_R2 = HabS2.LCL3450_L3_R2,JalS1_R1 = JalS1.LCL3442_L3_R1,JalS1_R2 = JalS1.LCL3442_L3_R2,CayP2_R1 = CayP2.LCL3431_L3_R1,CayP2_R2 = CayP2.LCL3431_L3_R2,CayP3_R1 = CayP3.LCL3433_L3_R1,CayP3_R2 = CayP3.LCL3433_L3_R2,CayS1_R1 = CayS1.LCL3430_L3_R1,CayS1_R2 = CayS1.LCL3430_L3_R2,CayS3_R1 = CayS3.LCL3432_L3_R1,CayS3_R2 = CayS3.LCL3432_L3_R2,SerP2_R1 = SerP2.LCL3447_L3_R1,SerP2_R2 = SerP2.LCL3447_L3_R2,SerS1_R1 = SerS1.LCL3444_L3_R1,SerS1_R2 = SerS1.LCL3444_L3_R2,SerS2_R1 = SerS2.LCL3446_L3_R1,SerS2_R2 = SerS2.LCL3446_L3_R2,WaxP2_R1 = WaxP2.LCL3441_L3_R1,WaxP2_R2 = WaxP2.LCL3441_L3_R2,WaxS1_R1 = WaxS1.LCL3438_L3_R1,WaxS1_R2 = WaxS1.LCL3438_L3_R2,WaxS2_R1 = WaxS2.LCL3440_L3_R1,WaxS2_R2 = WaxS2.LCL3440_L3_R2)


```

##Creating bar graph of Mapped Libraries
  Creating a bar graph to visualize the different size of mapped reads across the samples

```{r Bar Graph of Mapped Reads}

#The code kept getting held up so i commented it out momentarily

# trimmed_tidy %>%
# filter(Direction == "Forward") %>% 
#   view() 
# #Removes duplicate libraries
# ggplot(aes(x = Sample_ID, y = Mapped_Reads, fill = Pepper_Name)) +
# geom_col() + theme_bw() +
# scale_fill_brewer(palette = "Accent") +
# theme(axis.text.x = element_text(angle = 45,
#                                  hjust = 1)) +
# geom_text(aes(label = Mapped_Reads)) +
# labs(x = "Sample ID",
#      y = "Mapped Reads",
#      fill = "Pepper Name",
#      title = "Mapped Reads per Sample")


```


##Creating Box Plot of FPKM values
##Creating a box plot to visualize the FPKM values plotted against their Sample IDs

```{r Box Plot of FPKM values}

trimmed_tidy %>%
  select(Sample_ID, Direction, FPKM, Pepper_Name) %>%
  mutate(FPKM = ifelse(FPKM == 0, 1, FPKM), 
         FPKM_log2 = log2(FPKM)) %>%
  mutate(Direction = case_when(Direction == "Forward" ~ "R1",
                               Direction == "Reverse" ~ "R2")) %>%
  unite(Full_ID, Sample_ID:Direction, sep = "_") %>%
  ggplot(aes(x = Full_ID,
             y = FPKM_log2,
             fill = Pepper_Name,
             color = Pepper_Name,)) +
  geom_boxplot(alpha = 0.5) +
  theme_bw() +
  scale_fill_brewer(palette = "Dark2",              #Adding color to the graphs      
                    guide = "none") +
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x = element_text(angle = 45,      #Adjusting text location
                                   hjust = 1)) +
  labs(x = "Sample ID",                             #Adding labels
       y = "FPKM",
       color = "Pepper Name",
       title = "Log2 Transformartion of FPKM per Sample")
  

```

##Creating PCAs for Pepper Samples

```{r Attempting to Create a Naming Function, echo = FALSE}

splom_object %>%
  select(-Gene_ID) %>%
  colnames() -> pepper_columns

naming_function <- function()
{

  if(str_detect(pepper_columns, "Hab")){
  print("Habanero")
} else if (str_detect(pepper_columns, "Jal")){
  print("Jalapeno")
} else if (str_detect(pepper_columns, "Ser")){
  print("Serrano")
} else if (str_detect(pepper_columns, "Wax")){
  print("Hungarian Wax")
} else if (str_detect(pepper_columns, "Che")){
  print("Cherry")
} else if (str_detect(pepper_columns, "Gho")){
  print("Ghost")
} else {print("Cayenne")}
}

```



```{r PCA of Pepper Samples}
# sample_pca <- genes_fpkm %>%
#   pivot_longer(cols = 4:51,
#                names_to = "Sample_ID",
#                values_to = "FPKM") %>%
#   mutate(FPKM = ifelse(FPKM == 0, 1, FPKM), 
#          FPKM_log2 = log2(FPKM)) %>%
#   mutate(Sample_ID = str_sub(Sample_ID, 1, -5)) %>%
#   rename(Gene_ID = genes_fpkm_with_protein_charact) %>%
#   select(c(-Protein_Function, -locus, -FPKM)) %>%
#   pivot_wider(names_from = Sample_ID,
#               values_from = FPKM_log2) %>%
#   select(-Gene_ID) %>%
#   scale() %>%
#   prcomp()
# 
# 
# #### Using dplyr to manipulate the data we have into a format we want ####
# 
# sample_identity <- final_tidy %>% select(Sample_ID, Pepper_Name, Tissue) %>% unique()
# 
# pre_pca <- genes_fpkm %>%
#   pivot_longer(cols = 4:51,
#                names_to = "Sample_ID",
#                values_to = "FPKM") %>%
#   mutate(FPKM = ifelse(FPKM == 0, 1, FPKM), 
#          FPKM_log2 = log2(FPKM)) %>%
#   rename(Gene_ID = genes_fpkm_with_protein_charact) %>%
#   select(c(-Protein_Function, -locus, -FPKM)) %>%
#   pivot_wider(names_from = Gene_ID,
#               values_from = FPKM_log2) %>%
#   left_join(sample_identity) %>%
#   mutate(Sample_ID = str_sub(Sample_ID, 1, -5))
# 
# #### Running the PCA ####
# pre_pca %>%
#   select(-Sample_ID, -Pepper_Name, -Tissue) %>%
#   scale() %>% #Normalizes data for clustering
#   prcomp() -> pepper_pca
# 
# #### Plotting PC1 v PC2 ####
# as_tibble(pepper_pca$x) %>% #Generates table
#   mutate(Sample_ID = pre_pca$Sample_ID,
#          Pepper_Name = pre_pca$Pepper_Name) %>%
#   ggplot(aes(x = PC1, y = PC2, color = Pepper_Name)) +
#   geom_point(size = 1,
#              alpha=0.5) +
#   theme_bw()
# 
# #### Plotting PC1 v PC3 ####
# as_tibble(pepper_pca$x) %>% #Generates table
#   mutate(Tissue = ifelse(str_detect(Sample_ID, "P"), "Placenta", "Skin")) %>%
#   mutate(Direction = ifelse(str_detect(Sample_ID, "R1"), "Forward", "Reverse")) %>%
#   ggplot(aes(x = PC1, y = PC3, color = Tissue)) +
#   geom_point() + theme_bw()
# 
# #### Plotting PC2 v PC3 ####
# as_tibble(pepper_pca$rotation) %>% #Generates table
#   mutate(Tissue = ifelse(str_detect(Sample_ID, "P"), "Placenta", "Skin")) %>%
#   mutate(Direction = ifelse(str_detect(Sample_ID, "R1"), "Forward", "Reverse")) %>%
#   ggplot(aes(x = PC2, y = PC3, color = Direction)) +
#   geom_point() + theme_bw()
# 
#   
# #### Creating a Scree Plot to show PC variation ####
#   var_explained = pepper_pca$sdev^2 / sum(pepper_pca$sdev^2) #Defining the amount of variation explained by each PC
# 
# 
#   qplot(c(1:48), var_explained) + #Using qplot to quickly select info out of a value
#   geom_line() + geom_text(aes(label = pepper_pca$x)) +
#   xlab("Principal Component") + 
#   ylab("Variance Explained") +
#   ggtitle("Scree Plot") +
#   ylim(0, 1) +
#   theme_bw()

```

##Creating Scatterplot of FPKM Values

Creating a scatter plot that plots the FPKM values for a forward run sample against its reverse run counterpart

```{r Scatterplot of FPKM values}

# Defining renaming variable to avoid chunky code

sample_cols <- c("HabP1_R1 = HabP1_R1,HabP1_R2 = HabP1_R2,JalP1_R1 = JalP1_R1, JalP1_R2 = JalP1_R2,SerP1_R1 = SerP1_R1,SerP1_R2 = SerP1_R2,WaxP1_R1 = WaxP1_R1,WaxP1_R2 = WaxP1_R2,CheP1_R1 = CheP1.LCL3435_L3_R1,CheP1_R2 = CheP1.LCL3435_L3_R2,CheP2_R1 = CheP2.LCL3437_L3_R1,CheP2_R2 = CheP2.LCL3437_L3_R2,CheS1_R1 = CheS1.LCL3434_L3_R1,CheS1_R2 = CheS1.LCL3434_L3_R2,CheS2_R1 = CheS2.LCL3436_L3_R1,CheS2_R2 = CheS2.LCL3436_L3_R2,GhoP1_R1 = GhoP1.LCL3453_L3_R1,GhoP1_R2 = GhoP1.LCL3453_L3_R2,GhoS1_R1 = GhoS1.LCL3452_L3_R1,GhoS2_R2 = GhoS1.LCL3452_L3_R2,HabP2_R1 = HabP2.LCL3451_L3_R1,HabP2_R2 = HabP2.LCL3451_L3_R2,HabS1_R1 = HabS1.LCL3448_L3_R1,HabS1_R2 = HabS1.LCL3448_L3_R2,HabS2_R1 = HabS2.LCL3450_L3_R1,HabS2_R2 = HabS2.LCL3450_L3_R2,JalS1_R1 = JalS1.LCL3442_L3_R1,JalS1_R2 = JalS1.LCL3442_L3_R2,CayP2_R1 = CayP2.LCL3431_L3_R1,CayP2_R2 = CayP2.LCL3431_L3_R2,CayP3_R1 = CayP3.LCL3433_L3_R1,CayP3_R2 = CayP3.LCL3433_L3_R2,CayS1_R1 = CayS1.LCL3430_L3_R1,CayS1_R2 = CayS1.LCL3430_L3_R2,CayS3_R1 = CayS3.LCL3432_L3_R1,CayS3_R2 = CayS3.LCL3432_L3_R2,SerP2_R1 = SerP2.LCL3447_L3_R1,SerP2_R2 = SerP2.LCL3447_L3_R2,SerS1_R1 = SerS1.LCL3444_L3_R1,SerS1_R2 = SerS1.LCL3444_L3_R2,SerS2_R1 = SerS2.LCL3446_L3_R1,SerS2_R2 = SerS2.LCL3446_L3_R2,WaxP2_R1 = WaxP2.LCL3441_L3_R1,WaxP2_R2 = WaxP2.LCL3441_L3_R2,WaxS1_R1 = WaxS1.LCL3438_L3_R1,WaxS1_R2 = WaxS1.LCL3438_L3_R2,WaxS2_R1 = WaxS2.LCL3440_L3_R1,WaxS2_R2 = WaxS2.LCL3440_L3_R2")


  # splom(splom_object[c(4,5,28,29)],
  #       par.settings =simpleTheme(col="magenta4"),
  #       upper.panel = panel.cor, #Removes upper scatterplot
  #       varname.cex = 0.95, #changes pepper name size
  #       axis.text.cex = 0.65, #changes axis size
  #       xlab= "Jalapeno FPKM values", #x-axis label
  #       ylab= "Jalapeno FPKM values", #y-axis label
  #       pscales = 5, #number of equally spaced ticks on plot
  #       axis.line.lty = 3,
  #       px.scatter = TRUE,
  #       horizontal=TRUE,
  #       pch = ".")

#Creates Jalapeno Scatterplot
splom_object %>%
  select(starts_with("Jal")) %>% #Identifies pepper sample by name
  pairs(upper.panel = panel.cor, #Replaces upper panel scatterplots with r-squared values
        font.labels = 2, #changes sample label size
        main = "Jalapeno Pepper FPKM values", #graph title
        pch = ".", #graph point size
        col = "palegreen2") #changes point color

#Creates Serrrano Scatterplot
splom_object %>%
  select(starts_with("Ser")) %>% #Identifies pepper sample by name
  pairs(upper.panel = panel.cor, #Replaces upper panel scatterplots with r-squared values
        font.labels = 2, #changes sample label size
        main = "Serrano Pepper FPKM values", #graph title
        pch = ".", #graph point size
        col = "peachpuff") #changes point color

#Creates Habanero Scatterplot
splom_object %>%
  select(starts_with("Hab")) %>% #Identifies pepper sample by name
  pairs(upper.panel = panel.cor, #Replaces upper panel scatterplots with r-squared values
        font.labels = 2, #changes sample label size
        main = "Habanero Pepper FPKM values", #graph title
        pch = ".", #graph point size
        col = "darkorange") #changes point color

#Creates Hungarian Wax Scatterplot
splom_object %>%
  select(starts_with("Wax")) %>% #Identifies pepper sample by name
  pairs(upper.panel = panel.cor, #Replaces upper panel scatterplots with r-squared values
        font.labels = 2, #changes sample label size
        main = "Hungarian Wax Pepper FPKM values", #graph title
        pch = ".", #graph point size
        col = "darkolivegreen") #changes point color

#Creates Cayenne Scatterplot
splom_object %>%
  select(starts_with("Cay")) %>% #Identifies pepper sample by name
  pairs(upper.panel = panel.cor, #Replaces upper panel scatterplots with r-squared values
        font.labels = 2, #changes sample label size
        main = "Cayenne Pepper FPKM values", #graph title
        pch = ".", #graph point size
        col = "gold1") #changes point color

#Creates Ghost Scatterplot
splom_object %>%
  select(starts_with("Gho")) %>% #Identifies pepper sample by name
  pairs(upper.panel = panel.cor, #Replaces upper panel scatterplots with r-squared values
        font.labels = 2, #changes sample label size
        main = "Ghost Pepper FPKM values", #graph title
        pch = ".", #graph point size
        col = "firebrick2") #changes point color

#Creates Cherry Scatterplot
splom_object %>%
  select(starts_with("Che")) %>% #Identifies pepper sample by name
  pairs(upper.panel = panel.cor, #Replaces upper panel scatterplots with r-squared values
        font.labels = 2, #changes sample label size
        main = "Cherry Pepper FPKM values", #graph title
        pch = ".", #graph point size
        col = "chartreuse4") #changes point color

```

##Creating HeatMap for Pepper samples

Creating a heatmap to show the top highly expressed genes utilizing their FPKM values

```{r HeatMap of FPKM values}
# genes_fpkm %>%
#   pivot_longer(cols = 4:51,
#                names_to = "Sample_ID",
#                values_to = "FPKM") %>%
#   mutate(FPKM = ifelse(FPKM == 0, 1, FPKM), 
#          FPKM_log2 = log2(FPKM)) %>%
#   mutate(Sample_ID = str_sub(Sample_ID, 1, -5)) %>%
#   rename(Gene_ID = genes_fpkm_with_protein_charact) %>%
#   select(c(-Protein_Function, -locus, -FPKM)) %>%
#   pivot_wider(names_from = Sample_ID,
#               values_from = FPKM_log2) %>%
#   select(-Gene_ID) %>%
#   heatmap(as.matrix())

```


###Corrplot for Pepper Samples

```{r Corrplot}

#Creating Ryan's color palettes for the scale

#ryan_pal <- colorRampPalette(c("#7FD1B9", "#AAE5D3", "#CEF0E6", "#E7F6F7"))
#ryan_pal2 <- colorRampPalette(c("#E7F6F7", "#CEF0E6", "#AAE5D3", "#7FD1B9"))
#ryan_pal3 <- colorRampPalette(c( "#84dcc6","#d6edff", "#8b95c9","#478978"))
ryan_pal4 <- colorRampPalette(c("#FE5F55", "#ECABA8", "#D9F7FA", "#2292A4"))
ryan_pal8 <- colorRampPalette(c("#2292A4", "#D9F7FA", "#ECABA8", "#FE5F55"))
ryan_pal5 <- colorRampPalette(c("#FE5F55", "#D9F7FA", "#236A80", "#2292A4"))

splom_object %>%
  select(-Gene_ID) %>%
  cor() %>%
  corrplot(order = "hclust", #Specifies how we arrange data
           method = "square", #Presentation of clusters
           addrect = 8, #Distinguishes groups among clusters
           rect.col = "blue3", #Color of groups
           tl.cex = 0.75, #Label size
           tl.col = "black", #Label colors
           tl.srt = 90, #Angle of labels - TOP
           bg = "white", #Background color
           cl.lim = c(0.5,1), #Range of the scale
           cl.ratio = 0.1, #Scale size
           cl.cex = 1, #Scale number size
           is.corr = F, #Whether the range affects the chosen colors
           col = ryan_pal4(5)) #Specifies color palette


#ryan_pal6 <- colorRampPalette(c("#FF9B71", "#FFFD82", "#9BBDF9", "#ED217C"))
#ryan_pal7 <- colorRampPalette(c("#F26430", "#F6FEDB", "#7678ED", "#F26430"))

#Removing wax placenta samples as they seem to interfere with data
splom_object %>%
  select(-c(Gene_ID, WaxP1_R1, WaxP1_R2)) %>%
  cor() %>%
  corrplot(order = "hclust", #Specifies how we arrange data
           method = "square", #Presentation of clusters
           addrect = 8, #Distinguishes groups among clusters
           rect.col = "blue3", #Color of groups
           tl.cex = 0.75, #Label size
           tl.col = "black", #Label colors
           tl.srt = 90, #Angle of labels - TOP
           bg = "white", #Background color
           cl.lim = c(0.5,1), #Range of the scale
           cl.ratio = 0.1, #Scale size
           cl.cex = 1, #Scale number size
           is.corr = F, #Whether the range affects the chosen colors
           col = ryan_pal5(8)) #Specifies color palette
```


###MDSfor Pepper Samples
```{r MDS}

#Log2 Principal components analysis (PCA)
genes_fpkm %>%
  rename(HabP1_R1 = HabP1_R1_001,HabP1_R2 = HabP1_R2_001,JalP1_R1 = JalP1_R1_001, JalP1_R2 = JalP1_R2_001,SerP1_R1 = SerP1_R1_001,SerP1_R2 = SerP1_R2_001,WaxP1_R1 = WaxP1_R1_001,WaxP1_R2 = WaxP1_R2_001,CheP1_R1 = CheP1.LCL3435_L3_R1_001,CheP1_R2 = CheP1.LCL3435_L3_R2_001,CheP2_R1 = CheP2.LCL3437_L3_R1_001,CheP2_R2 = CheP2.LCL3437_L3_R2_001,CheS1_R1 = CheS1.LCL3434_L3_R1_001,CheS1_R2 = CheS1.LCL3434_L3_R2_001,CheS2_R1 = CheS2.LCL3436_L3_R1_001,CheS2_R2 = CheS2.LCL3436_L3_R2_001,GhoP1_R1 = GhoP1.LCL3453_L3_R1_001,GhoP1_R2 = GhoP1.LCL3453_L3_R2_001,GhoS1_R1 = GhoS1.LCL3452_L3_R1_001,GhoS2_R2 = GhoS1.LCL3452_L3_R2_001,HabP2_R1 = HabP2.LCL3451_L3_R1_001,HabP2_R2 = HabP2.LCL3451_L3_R2_001,HabS1_R1 = HabS1.LCL3448_L3_R1_001,HabS1_R2 = HabS1.LCL3448_L3_R2_001,HabS2_R1 = HabS2.LCL3450_L3_R1_001,HabS2_R2 = HabS2.LCL3450_L3_R2_001,JalS1_R1 = JalS1.LCL3442_L3_R1_001,JalS1_R2 = JalS1.LCL3442_L3_R2_001,CayP2_R1 = CayP2.LCL3431_L3_R1_001,CayP2_R2 = CayP2.LCL3431_L3_R2_001,CayP3_R1 = CayP3.LCL3433_L3_R1_001,CayP3_R2 = CayP3.LCL3433_L3_R2_001,CayS1_R1 = CayS1.LCL3430_L3_R1_001,CayS1_R2 = CayS1.LCL3430_L3_R2_001,CayS3_R1 = CayS3.LCL3432_L3_R1_001,CayS3_R2 = CayS3.LCL3432_L3_R2_001,SerP2_R1 = SerP2.LCL3447_L3_R1_001,SerP2_R2 = SerP2.LCL3447_L3_R2_001,SerS1_R1 = SerS1.LCL3444_L3_R1_001,SerS1_R2 = SerS1.LCL3444_L3_R2_001,SerS2_R1 = SerS2.LCL3446_L3_R1_001,SerS2_R2 = SerS2.LCL3446_L3_R2_001,WaxP2_R1 = WaxP2.LCL3441_L3_R1_001,WaxP2_R2 = WaxP2.LCL3441_L3_R2_001,WaxS1_R1 = WaxS1.LCL3438_L3_R1_001,WaxS1_R2 = WaxS1.LCL3438_L3_R2_001,WaxS2_R1 = WaxS2.LCL3440_L3_R1_001,WaxS2_R2 = WaxS2.LCL3440_L3_R2_001) %>% #Renaming the peppernames to their shortened abbreviations
  pivot_longer(cols = 4:51, 
               names_to = "Sample_ID",
               values_to = "FPKM") %>% #Pivoting the data to then change to log2
  mutate(FPKM = ifelse(FPKM == 0, 1, FPKM), #Changing the 0s to 1s so log2 doesnt end up at negative infinity
         FPKM_log2 = log2(FPKM)) %>% #log2 function
  rename(Gene_ID = genes_fpkm_with_protein_charact) %>% #renaming columns
  select(c(-Protein_Function, -locus, -FPKM)) %>% #selecting columns
  pivot_wider(names_from = Sample_ID, #pivoting back to workable data frame for PCA
              values_from = FPKM_log2) %>%
  select(-Gene_ID) %>% #removing the geneID column
  plotMDS(gene.selection= "common") #selects the same genes for all comparisons
          

#Principal components analysis (PCA)
genes_fpkm %>%
  rename(HabP1_R1 = HabP1_R1_001,HabP1_R2 = HabP1_R2_001,JalP1_R1 = JalP1_R1_001, JalP1_R2 = JalP1_R2_001,SerP1_R1 = SerP1_R1_001,SerP1_R2 = SerP1_R2_001,WaxP1_R1 = WaxP1_R1_001,WaxP1_R2 = WaxP1_R2_001,CheP1_R1 = CheP1.LCL3435_L3_R1_001,CheP1_R2 = CheP1.LCL3435_L3_R2_001,CheP2_R1 = CheP2.LCL3437_L3_R1_001,CheP2_R2 = CheP2.LCL3437_L3_R2_001,CheS1_R1 = CheS1.LCL3434_L3_R1_001,CheS1_R2 = CheS1.LCL3434_L3_R2_001,CheS2_R1 = CheS2.LCL3436_L3_R1_001,CheS2_R2 = CheS2.LCL3436_L3_R2_001,GhoP1_R1 = GhoP1.LCL3453_L3_R1_001,GhoP1_R2 = GhoP1.LCL3453_L3_R2_001,GhoS1_R1 = GhoS1.LCL3452_L3_R1_001,GhoS2_R2 = GhoS1.LCL3452_L3_R2_001,HabP2_R1 = HabP2.LCL3451_L3_R1_001,HabP2_R2 = HabP2.LCL3451_L3_R2_001,HabS1_R1 = HabS1.LCL3448_L3_R1_001,HabS1_R2 = HabS1.LCL3448_L3_R2_001,HabS2_R1 = HabS2.LCL3450_L3_R1_001,HabS2_R2 = HabS2.LCL3450_L3_R2_001,JalS1_R1 = JalS1.LCL3442_L3_R1_001,JalS1_R2 = JalS1.LCL3442_L3_R2_001,CayP2_R1 = CayP2.LCL3431_L3_R1_001,CayP2_R2 = CayP2.LCL3431_L3_R2_001,CayP3_R1 = CayP3.LCL3433_L3_R1_001,CayP3_R2 = CayP3.LCL3433_L3_R2_001,CayS1_R1 = CayS1.LCL3430_L3_R1_001,CayS1_R2 = CayS1.LCL3430_L3_R2_001,CayS3_R1 = CayS3.LCL3432_L3_R1_001,CayS3_R2 = CayS3.LCL3432_L3_R2_001,SerP2_R1 = SerP2.LCL3447_L3_R1_001,SerP2_R2 = SerP2.LCL3447_L3_R2_001,SerS1_R1 = SerS1.LCL3444_L3_R1_001,SerS1_R2 = SerS1.LCL3444_L3_R2_001,SerS2_R1 = SerS2.LCL3446_L3_R1_001,SerS2_R2 = SerS2.LCL3446_L3_R2_001,WaxP2_R1 = WaxP2.LCL3441_L3_R1_001,WaxP2_R2 = WaxP2.LCL3441_L3_R2_001,WaxS1_R1 = WaxS1.LCL3438_L3_R1_001,WaxS1_R2 = WaxS1.LCL3438_L3_R2_001,WaxS2_R1 = WaxS2.LCL3440_L3_R1_001,WaxS2_R2 = WaxS2.LCL3440_L3_R2_001) %>% #Renaming the peppernames to their shortened abbreviations
  select(-genes_fpkm_with_protein_charact, -Protein_Function, -locus) %>%
  plotMDS(gene.selection= "common", #selects the same genes for all comparisons
          #pch = 5, #changing plotting points
          cex = 1) #changing size of points
          
          

#Log2 Principal components analysis (PCoA = LogFC dim)
genes_fpkm %>%
  rename(HabP1_R1 = HabP1_R1_001,HabP1_R2 = HabP1_R2_001,JalP1_R1 = JalP1_R1_001, JalP1_R2 = JalP1_R2_001,SerP1_R1 = SerP1_R1_001,SerP1_R2 = SerP1_R2_001,WaxP1_R1 = WaxP1_R1_001,WaxP1_R2 = WaxP1_R2_001,CheP1_R1 = CheP1.LCL3435_L3_R1_001,CheP1_R2 = CheP1.LCL3435_L3_R2_001,CheP2_R1 = CheP2.LCL3437_L3_R1_001,CheP2_R2 = CheP2.LCL3437_L3_R2_001,CheS1_R1 = CheS1.LCL3434_L3_R1_001,CheS1_R2 = CheS1.LCL3434_L3_R2_001,CheS2_R1 = CheS2.LCL3436_L3_R1_001,CheS2_R2 = CheS2.LCL3436_L3_R2_001,GhoP1_R1 = GhoP1.LCL3453_L3_R1_001,GhoP1_R2 = GhoP1.LCL3453_L3_R2_001,GhoS1_R1 = GhoS1.LCL3452_L3_R1_001,GhoS2_R2 = GhoS1.LCL3452_L3_R2_001,HabP2_R1 = HabP2.LCL3451_L3_R1_001,HabP2_R2 = HabP2.LCL3451_L3_R2_001,HabS1_R1 = HabS1.LCL3448_L3_R1_001,HabS1_R2 = HabS1.LCL3448_L3_R2_001,HabS2_R1 = HabS2.LCL3450_L3_R1_001,HabS2_R2 = HabS2.LCL3450_L3_R2_001,JalS1_R1 = JalS1.LCL3442_L3_R1_001,JalS1_R2 = JalS1.LCL3442_L3_R2_001,CayP2_R1 = CayP2.LCL3431_L3_R1_001,CayP2_R2 = CayP2.LCL3431_L3_R2_001,CayP3_R1 = CayP3.LCL3433_L3_R1_001,CayP3_R2 = CayP3.LCL3433_L3_R2_001,CayS1_R1 = CayS1.LCL3430_L3_R1_001,CayS1_R2 = CayS1.LCL3430_L3_R2_001,CayS3_R1 = CayS3.LCL3432_L3_R1_001,CayS3_R2 = CayS3.LCL3432_L3_R2_001,SerP2_R1 = SerP2.LCL3447_L3_R1_001,SerP2_R2 = SerP2.LCL3447_L3_R2_001,SerS1_R1 = SerS1.LCL3444_L3_R1_001,SerS1_R2 = SerS1.LCL3444_L3_R2_001,SerS2_R1 = SerS2.LCL3446_L3_R1_001,SerS2_R2 = SerS2.LCL3446_L3_R2_001,WaxP2_R1 = WaxP2.LCL3441_L3_R1_001,WaxP2_R2 = WaxP2.LCL3441_L3_R2_001,WaxS1_R1 = WaxS1.LCL3438_L3_R1_001,WaxS1_R2 = WaxS1.LCL3438_L3_R2_001,WaxS2_R1 = WaxS2.LCL3440_L3_R1_001,WaxS2_R2 = WaxS2.LCL3440_L3_R2_001) %>% #Renaming the peppernames to their shortened abbreviations
  pivot_longer(cols = 4:51, 
               names_to = "Sample_ID",
               values_to = "FPKM") %>% #Pivoting the data to then change to log2
  mutate(FPKM = ifelse(FPKM == 0, 1, FPKM), #Changing the 0s to 1s so log2 doesnt end up at negative infinity
         FPKM_log2 = log2(FPKM)) %>% #log2 function
  rename(Gene_ID = genes_fpkm_with_protein_charact) %>% #renaming columns
  select(c(-Protein_Function, -locus, -FPKM)) %>% #selecting columns
  pivot_wider(names_from = Sample_ID, #pivoting back to workable data frame for PCA
              values_from = FPKM_log2) %>%
  select(-Gene_ID) %>% #removing the geneID column
  plotMDS(gene.selection= "pairwise") #chooses the top genes separatley for each pariwise comparison between samples


  
#Principal components of Analysis (PCoA = LogFC dim)
genes_fpkm %>%
  rename(HabP1_R1 = HabP1_R1_001,HabP1_R2 = HabP1_R2_001,JalP1_R1 = JalP1_R1_001, JalP1_R2 = JalP1_R2_001,SerP1_R1 = SerP1_R1_001,SerP1_R2 = SerP1_R2_001,WaxP1_R1 = WaxP1_R1_001,WaxP1_R2 = WaxP1_R2_001,CheP1_R1 = CheP1.LCL3435_L3_R1_001,CheP1_R2 = CheP1.LCL3435_L3_R2_001,CheP2_R1 = CheP2.LCL3437_L3_R1_001,CheP2_R2 = CheP2.LCL3437_L3_R2_001,CheS1_R1 = CheS1.LCL3434_L3_R1_001,CheS1_R2 = CheS1.LCL3434_L3_R2_001,CheS2_R1 = CheS2.LCL3436_L3_R1_001,CheS2_R2 = CheS2.LCL3436_L3_R2_001,GhoP1_R1 = GhoP1.LCL3453_L3_R1_001,GhoP1_R2 = GhoP1.LCL3453_L3_R2_001,GhoS1_R1 = GhoS1.LCL3452_L3_R1_001,GhoS2_R2 = GhoS1.LCL3452_L3_R2_001,HabP2_R1 = HabP2.LCL3451_L3_R1_001,HabP2_R2 = HabP2.LCL3451_L3_R2_001,HabS1_R1 = HabS1.LCL3448_L3_R1_001,HabS1_R2 = HabS1.LCL3448_L3_R2_001,HabS2_R1 = HabS2.LCL3450_L3_R1_001,HabS2_R2 = HabS2.LCL3450_L3_R2_001,JalS1_R1 = JalS1.LCL3442_L3_R1_001,JalS1_R2 = JalS1.LCL3442_L3_R2_001,CayP2_R1 = CayP2.LCL3431_L3_R1_001,CayP2_R2 = CayP2.LCL3431_L3_R2_001,CayP3_R1 = CayP3.LCL3433_L3_R1_001,CayP3_R2 = CayP3.LCL3433_L3_R2_001,CayS1_R1 = CayS1.LCL3430_L3_R1_001,CayS1_R2 = CayS1.LCL3430_L3_R2_001,CayS3_R1 = CayS3.LCL3432_L3_R1_001,CayS3_R2 = CayS3.LCL3432_L3_R2_001,SerP2_R1 = SerP2.LCL3447_L3_R1_001,SerP2_R2 = SerP2.LCL3447_L3_R2_001,SerS1_R1 = SerS1.LCL3444_L3_R1_001,SerS1_R2 = SerS1.LCL3444_L3_R2_001,SerS2_R1 = SerS2.LCL3446_L3_R1_001,SerS2_R2 = SerS2.LCL3446_L3_R2_001,WaxP2_R1 = WaxP2.LCL3441_L3_R1_001,WaxP2_R2 = WaxP2.LCL3441_L3_R2_001,WaxS1_R1 = WaxS1.LCL3438_L3_R1_001,WaxS1_R2 = WaxS1.LCL3438_L3_R2_001,WaxS2_R1 = WaxS2.LCL3440_L3_R1_001,WaxS2_R2 = WaxS2.LCL3440_L3_R2_001) %>% #Renaming the peppernames to their shortened abbreviations
  select(-genes_fpkm_with_protein_charact, -Protein_Function, -locus) %>%
  plotMDS(gene.selection= "pairwise")



#Using Glimma package Principal components of Analysis (PCoA = LogFC dim)
genes_fpkm %>%
  rename(HabP1_R1 = HabP1_R1_001,HabP1_R2 = HabP1_R2_001,JalP1_R1 = JalP1_R1_001, JalP1_R2 = JalP1_R2_001,SerP1_R1 = SerP1_R1_001,SerP1_R2 = SerP1_R2_001,WaxP1_R1 = WaxP1_R1_001,WaxP1_R2 = WaxP1_R2_001,CheP1_R1 = CheP1.LCL3435_L3_R1_001,CheP1_R2 = CheP1.LCL3435_L3_R2_001,CheP2_R1 = CheP2.LCL3437_L3_R1_001,CheP2_R2 = CheP2.LCL3437_L3_R2_001,CheS1_R1 = CheS1.LCL3434_L3_R1_001,CheS1_R2 = CheS1.LCL3434_L3_R2_001,CheS2_R1 = CheS2.LCL3436_L3_R1_001,CheS2_R2 = CheS2.LCL3436_L3_R2_001,GhoP1_R1 = GhoP1.LCL3453_L3_R1_001,GhoP1_R2 = GhoP1.LCL3453_L3_R2_001,GhoS1_R1 = GhoS1.LCL3452_L3_R1_001,GhoS2_R2 = GhoS1.LCL3452_L3_R2_001,HabP2_R1 = HabP2.LCL3451_L3_R1_001,HabP2_R2 = HabP2.LCL3451_L3_R2_001,HabS1_R1 = HabS1.LCL3448_L3_R1_001,HabS1_R2 = HabS1.LCL3448_L3_R2_001,HabS2_R1 = HabS2.LCL3450_L3_R1_001,HabS2_R2 = HabS2.LCL3450_L3_R2_001,JalS1_R1 = JalS1.LCL3442_L3_R1_001,JalS1_R2 = JalS1.LCL3442_L3_R2_001,CayP2_R1 = CayP2.LCL3431_L3_R1_001,CayP2_R2 = CayP2.LCL3431_L3_R2_001,CayP3_R1 = CayP3.LCL3433_L3_R1_001,CayP3_R2 = CayP3.LCL3433_L3_R2_001,CayS1_R1 = CayS1.LCL3430_L3_R1_001,CayS1_R2 = CayS1.LCL3430_L3_R2_001,CayS3_R1 = CayS3.LCL3432_L3_R1_001,CayS3_R2 = CayS3.LCL3432_L3_R2_001,SerP2_R1 = SerP2.LCL3447_L3_R1_001,SerP2_R2 = SerP2.LCL3447_L3_R2_001,SerS1_R1 = SerS1.LCL3444_L3_R1_001,SerS1_R2 = SerS1.LCL3444_L3_R2_001,SerS2_R1 = SerS2.LCL3446_L3_R1_001,SerS2_R2 = SerS2.LCL3446_L3_R2_001,WaxP2_R1 = WaxP2.LCL3441_L3_R1_001,WaxP2_R2 = WaxP2.LCL3441_L3_R2_001,WaxS1_R1 = WaxS1.LCL3438_L3_R1_001,WaxS1_R2 = WaxS1.LCL3438_L3_R2_001,WaxS2_R1 = WaxS2.LCL3440_L3_R1_001,WaxS2_R2 = WaxS2.LCL3440_L3_R2_001) %>% #Renaming the peppernames to their shortened abbreviations
  select(-genes_fpkm_with_protein_charact, -Protein_Function, -locus) %>%
  glMDSPlot(gene.selection= "pairwise")


```
