---
title: "Pepper_Markdown"
author: "Ryan Calhoun"
date: "6/3/2021"
output: 
  html_document:
    theme: paper
    highlight: zenburn
    center: TRUE

editor_options: 
  chunk_output_type: console
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(RColorBrewer)

setwd("/Users/Caronelab/Desktop/Spicy_Pepper_Project/R_Workflow")
#setwd()

```

## Loading in Raw Data
  Here we imported **all** of the raw data before _tidying_ it. All functions were performed utilizing the default **Tidyverse** package!
  
```{r Loading in Data}

genes_fpkm <- read.csv ("genes_fpkm.csv") #Reading in FPKM Values
protein_data <- read.csv("Spicy_Pepper_cufflinks_information.csv") #Pepper information

```

## Tidying the Raw Data

```{r Tidying the Data}

tidy_fpkm <- genes_fpkm %>%
  pivot_longer(cols = 4:51,
               names_to = "Sample_ID",
               values_to = "FPKM") #Shrink dataframe width by creating 'Sample ID' col

final_tidy <- tidy_fpkm %>%
  full_join(protein_data,
             by = "Sample_ID") %>%
  rename(c(Locus = locus,
           Gene_ID = genes_fpkm_with_protein_charact)) 
#Add pepper information to tidied dataset, rename cols to make working w data easier

trimmed_tidy <- final_tidy %>%
  mutate(Sample_ID = str_sub(Sample_ID, 1, -5)) %>%
  mutate(Forward = str_detect(Sample_ID, "R1")) %>%
  mutate(Direction = case_when(Forward == TRUE ~ "Forward",
                               Forward == FALSE ~ "Reverse")) %>%
  select(-Forward) %>%
  mutate(Sample_ID = str_sub(Sample_ID,1, 5))

```

##Creating bar graph of Mapped Libraries
  Creating a bar graph to visualize the different size of mapped reads across the samples

```{r Bar Graph of Mapped Reads}

  trimmed_tidy %>%
  filter(Direction == "Forward") %>%
  ggplot(aes(x = Sample_ID, y = Mapped_Reads, fill = Pepper_Name)) +
  geom_col() + theme_bw() +
  scale_fill_brewer(palette = "Accent") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Sample ID", y = "Mapped Reads", fill = "Pepper Name", title = "Mapped Reads per Sample")
  

```


##Creating Box Plot of FPKM values

```{r Box Plot of FPKM values}
#Creating a box plot to visualize the FPKM values plotted against their Sample IDs

  


```

##Creating PCAs for Pepper Samples

```{r PCA of Pepper Samples}
#Creating a PCA plot for the FPKM values against the different sample IDs

#Creating a PCA plot for the skin vs placenta tissue

#Creating a PCA plot for the Capsicum annuum vs Capsicum chinense species

#Creating a PCA plot for the scoville scale for the different peppers




```

##Creating Scatterplot of FPKM Values

```{r Scatterplot of FPKM values}
#Creating a scatter plot that plots the FPKM values for a forward run sample against its reverse run counterpart







```

##Creating HeatMap for Pepper samples

```{r HeatMap of FPKM values}
#Creating a heatmap to show the top highly expressed genes utilizing their FPKM values






```