---
title: "Thesis Data Viz"
author: "Ryan Calhoun"
date: "6/3/2021"
output: 
  html_document:
    theme: paper
    highlight: zenburn
    center: TRUE

editor_options: 
  chunk_output_type: console
  
---

```{r setup, include=FALSE, fig.align = "center"}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(RColorBrewer)
library(lattice)


setwd("/Users/Caronelab/Desktop/Spicy_Pepper_Project/R_Workflow")
#setwd()

```

## Loading in Raw Data
  Here we imported **all** of the raw data before _tidying_ it. All functions were performed utilizing the default **Tidyverse** package!
  
```{r Loading in Data}

genes_fpkm <- read.csv ("genes_fpkm.csv") #Reading in FPKM Values
protein_data <- read.csv("Spicy_Pepper_cufflinks_information.csv") #Pepper information
splom_data <- read.csv("splom.csv")

```

## Tidying the Raw Data

```{r Tidying the Data}

tidy_fpkm <- genes_fpkm %>%
  pivot_longer(cols = 4:51,
               names_to = "Sample_ID",
               values_to = "FPKM") #Shrink dataframe width by creating 'Sample ID' col

final_tidy <- tidy_fpkm %>%
  full_join(protein_data,
             by = "Sample_ID") %>%
  rename(c(Locus = locus,
           Gene_ID = genes_fpkm_with_protein_charact)) 
#Add pepper information to tidied dataset, rename cols to make working w data easier

trimmed_tidy <- final_tidy %>%
  mutate(Sample_ID = str_sub(Sample_ID, 1, -5)) %>%
  mutate(Forward = str_detect(Sample_ID, "R1")) %>%
  mutate(Direction = case_when(Forward == TRUE ~ "Forward",
                               Forward == FALSE ~ "Reverse")) %>%
  select(-Forward) %>%
  mutate(Sample_ID = str_sub(Sample_ID,1, 5))

```

##Creating bar graph of Mapped Libraries
  Creating a bar graph to visualize the different size of mapped reads across the samples

```{r Bar Graph of Mapped Reads}

  trimmed_tidy %>%
  filter(Direction == "Forward") %>%
  ggplot(aes(x = Sample_ID, y = Mapped_Reads, fill = Pepper_Name)) +
  geom_col() + theme_bw() +
  scale_fill_brewer(palette = "Accent") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Sample ID", y = "Mapped Reads", fill = "Pepper Name", title = "Mapped Reads per Sample")
  

```


##Creating Box Plot of FPKM values
##Creating a box plot to visualize the FPKM values plotted against their Sample IDs

```{r Box Plot of FPKM values}

trimmed_tidy %>%
  select(Sample_ID, FPKM, Pepper_Name) %>%
  mutate(Log2_Trans = log2(FPKM)) %>%
  ggplot(aes(x=Sample_ID, y=Log2_Trans, fill = Pepper_Name, color = Pepper_Name)) +
  geom_boxplot(alpha = 0.5) +
  theme_bw() +
  scale_fill_brewer(palette = "Dark2", guide = "none") +
  scale_color_brewer(palette = "Dark2") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x = "Sample ID", y = "FPKM",
       fill = "Pepper Name",
       title = "Log2 Transformartion of FPKM per Sample")


```

##Creating PCAs for Pepper Samples

```{r PCA of Pepper Samples}
#Creating a PCA plot for the FPKM values against the different sample IDs

#Creating a PCA plot for the skin vs placenta tissue

#Creating a PCA plot for the Capsicum annuum vs Capsicum chinense species

#Creating a PCA plot for the scoville scale for the different peppers




```

##Creating Scatterplot of FPKM Values
##Creating a scatter plot that plots the FPKM values for a forward run sample against its reverse run counterpart

```{r Scatterplot of FPKM values}

log2_splom_data <- log2(splom_data[c(4:51)])

  Hab_scatter <- splom(log2_splom_data[c(1,2,21,22,23,24,25,26)])
  Jal_scatter <- splom(log2_splom_data[c(3,4,27,28)])
  Che_scatter <- splom(log2_splom_data[c(9,10,11,12,13,14,15,16)])
  Ser_scatter <- splom(log2_splom_data[c(5,6,37,38,39,40,41,42)])
  Gho_scatter <- splom(log2_splom_data[c(17,18,19,20)])
  Wax_scatter <- splom(log2_splom_data[c(7,8,43,44,45,46,47,48)])
  Cay_scatter <- splom(log2_splom_data[c(29,30,31,32,33,34,35,36)])

```

##Creating HeatMap for Pepper samples

```{r HeatMap of FPKM values}
#Creating a heatmap to show the top highly expressed genes utilizing their FPKM values

# tidy_fpkm %>%
#   pivot_wider(names_from = Sample_ID,
#               values_from = FPKM) %>%
#   select(-c(1:3)) %>%
#   scale() %>% #Normalizing the dataset
#   dist(method = "euclidean") %>% #Determining dist btw points
#   hclust(method = "average") %>% #Clustering by average, the default for h-clustering
#   plot()




```


